---
title: "Snapper-Grouper MSE: Preliminary Results"
subtitle: 'SAFMC Snapper-Grouper Advisory Panel'
author: "Adrian Hordyk"
institute: 'Blue Matter Science'
date: "`r Sys.Date()`"
header-includes: 
  - \titlegraphic{\includegraphics[width = 6cm]{logo.png}}
  - \AtBeginDocument{\title[Preliminary Results]{Snapper Grouper MSE\\Preliminary Results}}
output: 
  beamer_presentation:
    theme: "Berlin"
    colortheme: "default"
    fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction
## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

# Operating Models


# Management Options

# Performance Measures

# Results 
